---
title: "Course-Project-1"
output: html_document
---

The following initializes the necessary libraries for this code and reads the 
activity csv. 
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
activity <- read.csv("activity.csv")

```

## What is the mean and median total number of steps taken per day?
```{r}
sum_qd <- aggregate(steps ~ date, activity, sum)
hist(sum_qd$steps, xlab = "Number of Steps", main = "Total Number of Steps Per Day")
sq_mean <- mean(sum_qd$steps)
sq_median <- median(sum_qd$steps)
```

The mean number of steps per day is `r sq_mean` and the median number is `r sq_median`

## What is the average daily activity pattern?
```{r}
mean_qint <- aggregate(steps ~ interval, activity, mean)
plot(mean_qint, type = "l", xlab = "Interval", ylab = "Steps", main = "Average Steps per Interval")
max <- max(mean_qint$steps)
max_step_int <- filter(mean_qint, steps == max)
```

The interval with the highest average number of steps is `r max_step_int$interval`

## Inputing missing values
```{r}
total_na <- sum(is.na(activity$steps))
```

The number of missing values is `r total_na`. 

Missing values were replaced with average number of steps per interval
```{r}
activity$steps <- as.numeric(activity$steps)
activity1 <- mutate(activity, steps = if_else(is.na(steps), 
                                             mean(mean_qint$steps), steps))

sum_qd1 <- aggregate(steps ~ date, activity1, sum)
hist(sum_qd1$steps, xlab = "Number of Steps", main = "Total Number of Steps Per Day")
sq_mean1 <- mean(sum_qd1$steps)
sq_median1 <- median(sum_qd1$steps)
```

The mean number of steps per day is `r sq_mean1` and the median number is `r sq_median1`

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity1$date <- strptime(activity1$date, format = "%Y-%m-%d")
activity2 <- mutate(activity1, weekday = if_else(weekdays(date) == "Saturday" | weekdays(date) == "Sunday", "weekend", "weekday"))
wkday_data <- aggregate(steps ~ interval + weekday, activity2, mean)
par(mfrow = c(2,1))
with(filter(wkday_data, weekday == "weekday"), 
     plot(interval, steps, type = "l", main = "Weekday"))
with(filter(wkday_data, weekday == "weekend"), 
     plot(interval, steps, type = "l", main = "Weekend", xlab = "Interval", 
          ylab = "Number of Steps"))

```
